Turkey Hemorrhagic Enteritis Virus transcriptome profiling

### 2023-05-22 manuscript_thev_transcriptome.Rmd
Rerun entire analysis including mock switched mock fastq N1 with infected fastq S3 due to probable sample mixup or contamination of mock with virus.

---
output:
  word_document: default
  pdf_document: default
geometry: margin=1in
font-size: 11pt
header-includes:
  - \usepackage{helvet}
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left]{lineno}
  - \linenumbers
bibliography: transcriptome_refs.bib
csl: asm.csl
---

# Turkey Hemorrhagic Enteritis Virus transcriptome profiling

\vspace{10mm}

@@

inline_hook <- function(x) {
  paste0("***", x, "***")
}
knit_hooks$set(iniline = inline_hook)
```

@@

## RESULTS      
**Overview of sequencing data and analysis pipeline outputs**     
A previous study by Zeinab *et al* showed that most THEV transcripts were detectable beginning at 4 hours[@Aboezz2019]; hence, infected MDTC-RP19 cells were collected in triplicates at 4-, 12-, 24-, and 72-hours post-infection(h.p.i) to ensure an amply wide window for transcript capture. Our `r seq_len`bp paired-end RNA sequencing experiment yielded an average of `r tot_count` million total reads per timepoint and the reads mapping to the virus genome increased dramatically from `r cov_4hr` reads at 4 h.p.i to `r cov_72hr` million reads at 72 h.p.i. Altogether, `r total_mapped` million reads from all timepoints mapped to THEV genome providing good coverage/depth, leaving no regions unmapped.

@@

### RNA extraction and Sequencing   
Total RNA was extracted from infected cells using Thermofishers' RNAqueous™-4PCR Total RNA Isolation Kit `(#AM1914)` as per manufacturer's instructions. The RNA quantity and purity was initially assessed using nanodrop, and RNA was used only if the A260/A280 ratio was 2.0 ± 0.05 and the A260/A230 ratio was >2 and <2.2. Extracted total RNA samples were sent to `LC Sciences, Houston TX` for poly-A-tailed mRNA sequencing where RNA integrity was checked with `Agilent Technologies 2100 Bioanalyzer High Sensitivity DNA Chip` and poly(A) RNA sequencing library was prepared following `Illumina's TruSeq-stranded-mRNA sample preparation protocol`. Paired-end sequencing was performed on `Illumina's NovaSeq 6000 sequencing system`.

### Computational Analysis of RNA Sequencing Data: Mapping and Transcript characterization
Analysis of our sequencing reads were analyzed following a well established general outline described by Pertea _et al_ [@Pertea2016]. Briefly, sequencing reads were trimmed with the `FastQC`[@FastQC] program to acheive an overall `Mean Sequence Quality (Phred Score)` of 36. Trimmed reads were mapped to the complete sequence of avirulent turkey hemorrhagic enteritis virus strain Virginia (https://www.ncbi.nlm.nih.gov/nuccore/AY849321.1/) using `Hisat2` (version `xxxxxx`) [@Pertea2016] with default settings except increasing the CPU cores to 10 without relying on known splice sites. The generated `BAM` files from each infection timepoint were fed into `StringTie` [@Pertea2016] using a `gff3` file from NCBI containing the predicted genes of THEV as a guide. A custom script was used to consolidate all transcripts from all timepoints without redundancy, generating the final transcriptome of THEV.

@@

## FIGURES     
![Fig.1](results/r/thev_genomic_map.png)  
**Figure 1. _Genomic map of THEV avirulent strain_**. The central horizontal line represents the double-stranded DNA marked at 5kb intervals as white line breaks. Blocks represent viral genes. Blocks above the DNA line are transcribed rightward, those below are transcribed leftward. pTP, DBP and 33K predicted to be spliced are shown as having tails. Shaded regions indicates regions containing genes (colored red) that are not conserved across adenovirus genera (i.e. "genus-specific"). Genes colored in blue are "genus-common".  Gene colored in light green is conserved in all but Atadenoviruses. The U exon (light blue) is an incomplete gene present in almost all AdVs.

>>>>

---
title: Turkey Hemorrhagic Enteritis Virus transcriptome profiling
output:
  word_document: default
  pdf_document: default
geometry: margin=1in
font-size: 11pt
header-includes:
  - \usepackage{helvet}
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage[left]{lineno}
  - \linenumbers
bibliography: transcriptome_refs.bib
csl: asm.csl
editor_options:
  chunck_output_type: console
---
\vspace{10mm}

@@

inline_hook <- function(x) {
  paste0("***", x, "***")
}
knit_hooks$set(inline = inline_hook)
```

@@

## RESULTS      
**Overview of sequencing data and analysis pipeline outputs**     
A previous study by Zeinab *et al* showed that almost all THEV transcripts were detectable beginning at 4 hours[@Aboezz2019]; hence, infected MDTC-RP19 cells were collected in triplicates at 4-, 12-, 24-, and 72-hours post-infection(h.p.i) to ensure an amply wide window to sample all transcripts. Our `r seq_len` bp paired-end RNA sequencing experiment yielded an average of `r tot_count` million total reads per timepoint and the reads mapping to the virus genome increased dramatically from `r cov_4hr` reads at 4 h.p.i to `r cov_72hr` million reads at 72 h.p.i. Altogether, `r total_mapped` million reads from all timepoints mapped to THEV genome providing good coverage/depth, leaving no regions unmapped (__Table 1__).

@@

### RNA extraction and Sequencing   
Total RNA was extracted from infected cells using Thermofishers' RNAqueous™-4PCR Total RNA Isolation Kit `(#AM1914)` as per manufacturer's instructions. An agarose gel electrophoresis was performed to check RNA integrity. The RNA quantity and purity was initially assessed using nanodrop, and RNA was used only if the A260/A280 ratio was 2.0 ± 0.05 and the A260/A230 ratio was >2 and <2.2. Extracted total RNA samples were sent to `LC Sciences, Houston TX` for poly-A-tailed mRNA sequencing where RNA integrity was checked with `Agilent Technologies 2100 Bioanalyzer High Sensitivity DNA Chip` and poly(A) RNA sequencing library was prepared following `Illumina's TruSeq-stranded-mRNA sample preparation protocol`. Paired-end sequencing was performed on `Illumina's NovaSeq 6000 sequencing system`.

### Computational Analysis of RNA Sequencing Data: Mapping and Transcript characterization
Analysis of our sequencing reads were analyzed following a well established general outline described by Pertea _et al_ [@Pertea2016]. Briefly, sequencing reads were trimmed with the `FastQC - version 0.11.9` [@FastQC] program to acheive an overall `Mean Sequence Quality (Phred Score)` of 36. Trimmed reads were mapped to the complete sequence of avirulent turkey hemorrhagic enteritis virus strain Virginia (https://www.ncbi.nlm.nih.gov/nuccore/AY849321.1/) using `Hisat2 - version 2.2.1` [@Pertea2016] with default settings except increasing the CPU cores to 10 without relying on known splice sites. The generated `BAM` files from each infection timepoint were fed into `StringTie - version 2.2.1` [@Pertea2016] using a `gff3` file from NCBI containing the predicted genes of THEV as a guide. A custom script was used to consolidate all transcripts from all timepoints without redundancy, generating the final transcriptome of THEV.

@@

## TABLES AND FIGURES     
![Fig.1](results/r/thev_genomic_map.png)  
**Figure 1. _Genomic map of THEV avirulent strain_**. The central horizontal line represents the double-stranded DNA marked at 5kb intervals as white line breaks. Blocks represent viral genes. Blocks above the DNA line are transcribed rightward, those below are transcribed leftward. pTP, DBP and 33K predicted to be spliced are shown as having tails. Shaded regions indicates regions containing genes (colored red) that are not conserved across adenovirus genera (i.e. "genus-specific"). Genes colored in blue are "genus-common".  Gene colored in light green is conserved in all but Atadenoviruses. The U exon (light blue) is an incomplete gene present in almost all AdVs.
